###Loading the data
```{r}
file<-"C:/Users/gridA/Desktop/R course/reproducible research/activity.csv"
data<-read.csv(file, header=T, sep=",")
```

###What is mean total number of steps taken per day?

1.Histogram of the total number of steps taken each day
```{r}
data2 <- aggregate(steps ~ date, data = data, sum, na.rm = TRUE)

hist(data2$steps, breaks = 10, col = "blue", 
    main = "Histogram of Total Steps on Each Day", 
    xlab = "Total Steps of Each Day")
```

2.Calculate and report the mean and median total number of steps taken per day
```{r}
stepsmean <- mean(data2$steps, na.rm = T)
stepsmedian <- median(data2$steps, na.rm = T)
```
```{r}
print(stepsmean)
print(stepsmedian)
```

###What is the average daily activity pattern?

1.Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
data3 <- aggregate(steps ~ interval, data = data, mean, na.rm = TRUE)
plot(data3$interval, data3$steps, 
     type = "l", col = "red", 
   xlab = "Time Interval", ylab = "Averaged Steps", main = "Steps Averaged Across All Days")
```
	 
2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxstep <- data3[which.max(data3$steps), ]$interval
print(maxstep)
```

###Imputing missing values

1.Calculate and report the total number of missing values in the dataset
```{r}
missing_num <- sum(!complete.cases(data$steps))
print(missing_num)
```

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
dataFilled <- data
for (i in 1:nrow(dataFilled)) {
    if (is.na(dataFilled[i, 1])) 
        dataFilled[i, 1] <- data3[data3[, 1] == dataFilled[i, 3], ][, 2]
}
missing_num <- sum(is.na(dataFilled))
```

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
data_new <- aggregate(steps ~ date, data = dataFilled, sum, na.rm = TRUE)

hist(data_new$steps, breaks = 10, col = "green", 
    main = "Histogram of Total Steps on Each Day (Imputed)", 
    xlab = "Total Steps of Each Day")
	
mean_new <- mean(data_new$steps)
median_new <- median(data_new$steps)
```
```{r}
print(mean_new)
print(median_new)
```

###Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
dataFilled$weekday <- c("weekday")
dataFilled[weekdays(as.Date(dataFilled[, 2])) %in% c("Sunday", "Saturday"), 
    ][4] <- c("weekend")
dataFilled$weekday <- factor(dataFilled$weekday)
```

2.Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
data_new2 = aggregate(steps ~ interval + weekday, dataFilled, mean)
library(lattice)
xyplot(steps ~ interval | factor(weekday), data = data_new2, 
       type = "l", layout = c(1, 2), 
	   ylab = "Averaged steps", xlab="Interval",
       main = "Averaged steps for weekday days or weekend days")
```
